### 2015-06-23 chapters/ipa_meets_unicode.tex
edits

The International Phonetic Alphabet (IPA) is a common standard in linguistics to
transcribe sounds of spoken language into Latin-based characters. Although IPA
is reasonably easily adhered to with pen and paper, it is not trivial to encode
IPA characters electronically. Similar to the previous chapter, this chapter
will discuss various pitfalls with the Unicode encoding of the IPA.\@ We will
specifically refer to the 2005 version of the IPA \citep{IPA2005} and the 7.0 
version of Unicode \citep{Unicode2014}.

@@

There are a few pitfalls to be aware of when using the Unicode Standard to
encode IPA.\@ As we have said before, from a linguistic perspective it might
sometimes look like the Unicode Consortium is making incomprehensible decisions,
but it is important to realize that the consortium has tried and is continuing
to try to be as consistent as possible across a wide range of use cases, and it
does place linguistic traditions above other orthographic possibilities. In
general, we strongly suggest linguists not to complain about any decisions in
the Unicode Standard, but to try and understand the rationale of the Unicode
Consortium (which in our experience is almost always well-conceived) and devise
ways to work with any unexpected behavior. Many of the current problems derive
from the fact that the IPA is clearly historically based on the Latin script,
but different enough from most other Latin-based writing systems to warrant
special attention. This ambivalent status of the IPA glyphs (partly Latin,
partly special) is unfortunately also attested in the treatment of IPA in the
Unicode Standard. In retrospect, it might have been better to consider the IPA
(and other transcription systems) to be a special kind of script within
the Unicode Standard, and treat the obvious similarity to Latin glyphs as a
historical relic. All IPA glyphs would then have their own code points, instead
of the current situation in which some IPA glyphs have special code points,
while others are treated as being identical to the regular Latin characters.
Yet, the current situation, however unfortunate, is unlikely to change, so as
linguists we will have to learn to deal with the specific pitfalls of IPA within
the Unicode Standard. In this section, we will describe these pitfalls in some
detail.

@@

The ambivalent nature of IPA glyphs arises because, on the one hand, the IPA
uses Latin-based glyphs like <a>, <b> or <p>. From this perspective, the IPA
seems to be just another orthographic tradition using Latin characters, all of
which do not get a special treatment within the Unicode Standard (just like
e.g.~the French, German, or Danish orthographic traditions do not have a special
status). On the other hand, the IPA uses many special symbols (like turned <ɐ>,
mirrored <ɘ> and/or extended <ɧ> Latin glyphs) not found in any other Latin-based
writing system. For this reason a special block with code points, called
\textsc{IPA Extensions} was included already in the first version of the Unicode
Standard (Version 1.0 from 1991).

>>>>

The International Phonetic Alphabet (IPA) is a common standard in linguistics to
transcribe sounds of spoken language into Latin-based characters. Although IPA
is reasonably easily adhered to with pen and paper, it is not trivial to encode
IPA characters electronically. Similar to the previous chapter, this chapter
will discuss various pitfalls with the encoding of IPA in the Unicode Standard.\@ We will
specifically refer to the 2005 version of the IPA \citep{IPA2005} and the 7.0 
version of Unicode \citep{Unicode2014}.

@@

There are several pitfalls to be aware of when using the Unicode Standard to
encode IPA.\@ As we have said before, from a linguistic perspective it might
sometimes look like the Unicode Consortium is making incomprehensible decisions,
but it is important to realize that the consortium has tried and is continuing
to try to be as consistent as possible across a wide range of use cases, and it
does place linguistic traditions above other orthographic possibilities. In
general, we strongly suggest to linguists to not complain about any decisions in
the Unicode Standard, but to try and understand the rationale of the Unicode
Consortium (which in our experience is almost always well-conceived) and devise
ways to work with any unexpected behavior. Many of the current problems derive
from the fact that the IPA is clearly historically based on the Latin script,
but different enough from most other Latin-based writing systems to warrant
special attention. This ambivalent status of the IPA glyphs (partly Latin,
partly special) is unfortunately also attested in the treatment of IPA in the
Unicode Standard. In retrospect, it might have been better to consider the IPA
(and other transcription systems) to be a special kind of script within
the Unicode Standard, and treat the obvious similarity to Latin glyphs as a
historical relic. All IPA glyphs would then have their own code points, instead
of the current situation in which some IPA glyphs have special code points,
while others are treated as being identical to the regular Latin characters.
Yet, the current situation, however unfortunate, is unlikely to change, so as
linguists we will have to learn to deal with the specific pitfalls of IPA within
the Unicode Standard. In this section, we will describe these pitfalls in some
detail.

@@

The ambivalent nature of IPA glyphs arises because, on the one hand, the IPA
uses Latin-based glyphs like <a>, <b> or <p>. From this perspective, the IPA
seems to be just another orthographic tradition using Latin characters, all of
which do not get a special treatment within the Unicode Standard (just like
e.g.~the French, German, or Danish orthographic traditions do not have a special
status). On the other hand, the IPA uses many special symbols (like turned <ɐ>,
mirrored <ɘ> and/or extended <ɧ> Latin glyphs) not found in any other Latin-based
writing system. For this reason a special block with code points, called
\textsc{IPA Extensions} was already included in the first version of the Unicode
Standard (Version 1.0 from 1991).
